### 2010-12-03 report/report.tex
report update

\section{Abstract}
% A brief summary of the report. This is not an introduction -
% it should be complete. Include 1-2 sentences for each of the items
% below, up to and including conclusions.
This paper presents several approaches to optimizing a quadrupedal
trot gait for forward speed. Given a parameterized walk designed for
a quadruped robot designed and printed in the Cornell Computational
Synthesis Lab, we implement, test, and compare different learning strategies, 
including uniform and Gaussian random hill climbing, a form of policy 
gradient reinforcement learning\cite{1}, Simplex (Nelder-Mead), 
linear regression, SVM regression, and an evolutionary neural 
network (HyperNEAT)\cite{2}. Because the fastest learned walk was not
significantly faster than the fastest randomly generated walk, we
conjecture that the motion representation for the robot is more
integral to forward speed than the learning algorithm.

Keywords: Learning Control, Walking Robots, Multi Legged Robots

\section{Introduction}
% Motivate and abstractly describe the problem you are addressing and
% how you are addressing it. What is the problem? Why is it important?
% What is your basic approach? A short discussion of how it fits into
% related work in the area is also desirable. Summarize the basic results
% and conclusions that you will present.
We are using machine learning to design a gait for a quadruped
robot. The robot has an on-board computer, drivers for the motors, and
will have camera feedback. The code so far consists of the robot
class, optimization class, parameterized model, sine model, and camera
feedback.  Both the hardware and software are progressing on schedule.
Some promising gaits have been generated, and we anticipate a final
evaluation of the robots gait in November.

\section{Problem definition}
% Precisely define the problem you are addressing (i.e. formally specify
% the inputs and outputs). Elaborate on why this is an interesting and
% important problem.
We are testing several different learning methods to design a gait for 
a quadruped robot from the Cornell Computational Synthesis Lab. The metric 
for evaluation of the designed gait is speed. A comparison and evaluation 
of the many different methods available for optimizing the gait of legged 
robots will be useful for future work on this challenging multidimensional 
control problem.

\section{Method}
% Describe in reasonable detail the algorithm you are using to address this 
% problem. A pseudo-code description of the algorithm you are using is 
% frequently useful. If it makes sense for your project, trace through a 
% concrete example, showing how your algorithm processes this example. The 
% example should be complex enough to illustrate all of the important aspects 
% of the problem but simple enough to be easily understood. If possible, 
% an intuitively meaningful example is better than one with meaningless 
% symbols.
We implemented and tested 8 different learning strategies:
\begin{itemize}
\item \emph{Uniform random hill climbing}: Creates a random neighbor by
  randomly choosing one parameter to adjust, and changing it completely 
  randomly in UniformStrategy(). Then evaluates this neighbor by running 
  the robot with the newly chosen parameters. If this neighbor results in a
  longer distance walked then the previous best neighbor, we save this
  neighbor as the new best neighbor. We then repeat the whole process
  using the best neighbor as a base.
\item \emph{Gaussian random hill climbing}: Creates a random neighbor by
  changing each parameter randomly based on a normal distribution. Then
  evaluates this neighbor similarly to uniform random hill climbing.
\item \emph{N-dimensional policy gradient descent}: Estimates the policy
  gradient by evaluating \emph{t} randomly generated policies
  near an initial parameter vector. Then computes the average score for
  each parameter and adjusts the base policy according to the estimated
  gradient.
\item \emph{Random}: Randomly generates policies in the range of the motion
representation.
\item \emph{Simplex (Nelder-Mead) Method}: Creates a simplex with 6 vertices and tests the fitness of each vertex. The worst vertex is reflected over the centroid. If the reflected point is better than the second worst point and worse than the best point, then the reflected point replaces the worst. If the reflected point is better than the best point, the simplex is expanded in the direction of the reflected point. The better of the reflected and the expanded point replaces the worst point. If the reflected point is worse than the second worst point, then the simplex is contracted away from the reflected point. If the contracted point is better than the reflected point, the contracted point replaces the worst point. If the contracted point is worse than the reflected point, the entire simplex is shrunk.
\item \emph{Linear regression}: To initialize, chooses five random points
and trains on them using least-squares. Then in a loop, takes a fixed-size
step in the direction of the gradient, and retrains on all points so far.
\item \emph{SVM regression}
\item \emph{Evolutionary Neural Network (HyperNEAT)\cite{2}}
\end{itemize}

\section{Related work}
% Briefly explain who else worked on related problems in the past and what
% methods they used. Explain if you are using similar methods, or if your
% approach is different and if so - how (either is ok).

\section{System Architecture and Implementation}
% Describe how you implemented your system and how you structured it. 
% This should give an overview of the system, not a detailed 
% documentation of the code. The documentation of the code is part of 
% the code you hand in. You might want to comment on high-level design 
% decisions that you made. Also explain how you obtained your
% data and any pre-processing you did to it.

The quadruped robot has an on-board computer running Linux. The lower
level drivers are in C and the system is implemented in Python. Feedback about distance travelled is provided via an infrared LED and a Wii remote.

\begin{itemize}
\item \code{Robot} class: Class wrapper for commanding
  motion of the robot.  The \code{Robot} class takes care of the robot
  initialization, communication with the servos, and timing of the
  runs.  In addition, it prevents the servos from ever being commanded
  to a point outside their normal range (0 - 1023) as well as beyond
  points where limbs would collide with parts of the robot body.  The
  main class function, \code{run}, accepts a motion model (any
  function that takes a time argument and outputs a 9 dimensional
  position) and will run the robot using this motion model, including,
  if desired, smooth interpolation over time for the beginning and end
  of the run.

\item \code{Strategy} and \code{RunManager} class: The user has a choice between seven
  different learning strategies: uniform random hill climbing, Gaussian random
  hill climbing, N-dimensional policy gradient descent, random, 
  simplex, linear regression/prediction, and SVM regression/prediction.

\item \code{Parameterized Motion Model} abstract base class: Several functions exist, but none are within a class framework.

\item \code{SineModel} classes: Commands motors to positions
  based on a sine wave, creating a periodic pattern. Parameters are:
  amplitude, wavelength, scale inner vs outer motors, scale left vs right
  motors, scale back vs front motors. Currently a function, not a class.

\item Other derived motion model classes: Several versions of a SineModel
  exist, each with different parameters, though solely as functions.

\item \code{WiiTrackClient} and \code{WiiTrackServer} classes
 and hardware: A Wii remote tracks the location of the robot through an infrared LED mounted on top of the robot. We used CWiid library which is able to interface with the remote via bluetooth to determine the location of the infrared LED. This information is accessed through some functions we wrote and passed to the program. The program gets the robot's position at the beginning of each run and then again at the end. The program then calculates the net change in position and uses that as the distance walked.

\end{itemize}

\section{Experimental Evaluation}

\subsection{Methodology}
% What are the criteria you are using to evaluate your method? What
% specific hypotheses does your experiment test? Describe the 
% experimental methodology that you used. What are the dependent and 
% independent variables? For projects in machine learning, what is 
% the training/test data that was used, and why is it realistic or
% interesting? Exactly what performance data did you collect and how 
% are you presenting and analyzing it? Comparisons to competing methods 
% that address the same problem or to variations of your own algorithm 
% are particularly useful.
\begin{itemize}
\item Each algorithm was run on 3 different initial parameter vectors on the
physical robot. We stop each run after plateauing results (no improvement 
for one third of the policies seen so far). We evaluate each method based on the
amount of improvement seen from the initial parameter vectors, and on the
fastest speeds achieved during runs.
\end{itemize}

\subsection{Results}
% Present the quantitative results of your experiments. Graphical data
% presentation such as graphs and histograms are frequently better 
% than tables. Explain what are the basic differences revealed in the 
% data. If your method is stochastic or you analyze multiple datasets, 
% calculate mean and standard error for every quantitative
% performance result you show based on multiple runs.
\begin{itemize}
\item We have done runs of 
3 different inital parameter vectors for random hill climbing, 
gaussian random hill climbing, policy gradient descent, random,
simplex, and linear regression. We developed several
gaits that were about 4 times faster than the original hand-coded gait.

\item For vector A, linear regression worked significantly better than the other algorithms, resulting in a gait (27.58 body lengths/minute) that walked over twice as fast as the next best gait (uniform random hill climbing - 11.37 body lengths/minute).

\item Vector B resulted in similarly performing gaits with all the algorithms. Random produced the best gait (17.26 body lengths/minute) and uniform random hill climbing produced the worst (9.44 body lengths/minute). The remaining algorithms each produced a gait around 13 body lengths/minute.

\item For vector C, simplex and Gaussian random hill climbing each produced a gait that substantially outperformed the other algorithms. Simplex resulted in a gait of nearly 15 body lengths/minute  and Gaussian random hill climbing produced a gait of just over 13 body lengths/minute, whereas the other algorithms returned gaits of less than 5 body lengths/minute.

\item Based on the three trials, linear regression worked the best, followed by simplex. Gaussian random hill climbing performed about as well as random, but it is unclear whether Gaussian would continue to improve and become much better than random if allowed to run for more iterations. Uniform random hill climbing and policy gradient descent performed similarly.

\item We also coded \code{explore\_dimensions} to collect data for dimension
varying plots. We intend to do further explore this by running
\code{explore\_dimension} around a parameter that produces a medium-good gait.
\end{itemize}

\subsection{Discussion}
% Is your hypothesis supported? What conclusions do the results support
% about the strengths and weaknesses of your method compared to other 
% methods? How can the results be explained in terms of the underlying 
% properties of the algorithm and/or the data.
\begin{itemize}
\item  Since only three trials were tested with each algorithm and no algorithm consistently outperformed the others, there is a large standard error for each algorithm. For this reason, it is unclear if any algorithm is superior to another in this application. More trials with each algorithm would be necessary to reach a definitive ranking of the performance of the algorithms.

\item Each algorithm discovered at least one gait of over 10 body lengths/minute, including random. For this reason, we speculate that the motion model may be more critical than the learning algorithm.
\end{itemize}

\section{Future work}
% What are the major shortcomings of your current method? For each 
% shortcoming, propose additions or enhancements that would help overcome it.
\begin{itemize}
\item Different motion representations
\item Evolutionary algorithms/HyperNEAT\cite{2}
\item Learning how to turn
\item Open sourcing code for others to extend
\end{itemize}

\section{Conclusion}
% Briefly summarize the important results and conclusions presented in 
% the paper. What are the most important points illustrated by your 
% work? How will your results improve future research and applications 
% in the area?
Because the fastest learned walk was not
significantly faster than the fastest randomly generated walk, we
conjecture that the motion representation for the robot is more
integral to forward speed than the learning algorithm.

\section{References}
% A numbered list of all books, articles, and web pages you used in your 
% work or that you cited in your report. Each item must be cited from 
% somewhere within the text. When citing a URL, include date visited. 
% When citing a commercial product, list the company.

\section{Acknowledgments}
% List any people not on the team who helped you with your project: Your
% TA, other people you consulted with or had useful discussions (say in 
% a word or two what they did), people who proofread your report, and 
% any external code you used (libraries etc).
\begin{itemize}
\item Hod Lipson, Cornell Computational Synthesis Lab: advisor
\item Jim T\o rreson, University of Oslo: advisor
\item Juan Zagal, University of Chile: designed and printed robot and provided assistance
\item Jeff Clune, CCSL Red Couch: collaborated on HyperNEAT implementation/testing
\item Cooper Bills, Cornell University: assisted with Wii tracker development
\item Anshumali Srivastava, Cornell University: Teaching Assistant
\end{itemize}

\section{The Team}
% A table listing the names of the people who worked on the project, and who
% did what.
\begin{itemize}
\item Diana Hidalgo: 
\begin{itemize}
\item Simplex (Nelder-Mead)
\item \code{WiiTrackClient} and \code{WiiTrackServer} classes and hardware
\item Plotting and evaluating results
\end{itemize}
\item Sarah Nguyen
\begin{itemize}
\item Random
\item Uniform random hill climbing
\item Gaussian random hill climbing
\item N-dimensional policy gradient descent
\item Linear regression and prediction
\item \code{RunManager} and \code{Strategy} classes
\item Video editing
\end{itemize}
\item Jason Yosinski
\begin{itemize}
\item SVM regression and prediction
\item Evolutionary Neural Network (HyperNEAT)
\item \code{WiiTrackClient} and \code{WiiTrackServer} classes and hardware
\item \code{Robot}, \code{RunManager}, and \code{Strategy} classes
\item \code{Parameterized Motion Model} abstract base class
\item \code{SineModel} classes
\item \code{explore\_dimensions}
\item Plotting and evaluating results
\end{itemize}
\end{itemize}

\section{Appendix}
% Describe the code files you uploaded into CMS

A brief description of the code uploaded to the CMS follows:

\begin{itemize}
\item \code{optimize.py}: Determines a strategy to try and runs the robot
  with that strategy.
\item \code{RunManager.py}: Deals with all the details of running the robot,
  including choosing an initial parameter, running the robot multiple times, 
  tracking distance walked, and writing to the log file. Also includes the
  \code{explore\_dimensions} method.
\item \code{Strategy.py}: Contains all the different possible strategies, which
  will be passed as objects in \code{optimize.py}. 
\item \code{Robot.py}: Implements the \code{Robot} class, described in
  \secref{implement}.
\item \code{SineModel.py}: Implements a sine based motion model,
  described in \secref{implement}.
\item \code{Motion.py}: Motion helper functions.
\item \code{WiiTrackServer.py}: Broadcasts the position of the infrared LED>
\item \code{WiiTrackClient.py}: Connects to the WiiTrackServer to get the current position information.
\end{itemize}