### 2010-12-03 report/abstract.tex
Older changes, maybe too old?

We herein propose a system...
\edit{Write abstract}