### 2010-12-03 report/abstract.tex
Moved abstract to separate file, added ownstyles

% A brief summary of the report. This is not an introduction -
% it should be complete. Include 1-2 sentences for each of the items
% below, up to and including conclusions.
This paper presents several approaches to optimizing a quadrupedal
trot gait for forward speed. Given a parameterized walk designed for
a quadruped robot designed and printed in the Cornell Computational
Synthesis Lab, we implement, test, and compare different learning strategies, 
including uniform and Gaussian random hill climbing, a form of policy 
gradient reinforcement learning\cite{1}, Simplex (Nelder-Mead), 
linear regression, SVM regression, and an evolutionary neural 
network (HyperNEAT)\cite{2}. Because the fastest learned walk was not
significantly faster than the fastest randomly generated walk, we
conjecture that the motion representation for the robot is more
integral to forward speed than the learning algorithm.

Keywords: Learning Control, Walking Robots, Multi Legged Robots