### 2010-12-07 report/abstract.tex
much editing with sarah, diana

% A brief summary of the report. This is not an introduction -
% it should be complete. Include 1-2 sentences for each of the items
% below, up to and including conclusions.

This paper presents an array of approaches to optimizing a quadrupedal
gait for forward speed.  We implement, test, and compare different
learning strategies including uniform and Gaussian random hill
climbing, policy gradient reinforcement learning\cite{kohl},
Nelder-Mead simplex\cite{nm}, new predictive methods based on linear
and support vector regression, and an evolved neural network
(HyperNEAT)\cite{2}.  We compare results to a baseline random search
method.  Many of the methods resulted in walks significantly faster
than previously hand-tuned gaits.

%Because the fastest learned walk
%was not significantly faster than the fastest randomly generated walk,
%we conjecture that the motion representation for the robot is more
%integral to forward speed than the learning algorithm.

\edit{fix citations}

Keywords: Learning Control, Walking Robots, Gait Learning, Quadruped