### 2011-01-20 paper/acknowledgments.tex
Restructuring of report into multiple files to minimize conflicts, references working, dealt old sections to appropriate locations in paper framework.

\section{Acknowledgments}

Thanks to ..., some grant maybe, Jim T\o rreson, ...and the reviewers.

\edit{write this, old below}

% List any people not on the team who helped you with your project: Your
% TA, other people you consulted with or had useful discussions (say in 
% a word or two what they did), people who proofread your report, and 
% any external code you used (libraries etc).
\begin{itemize}
\item Hod Lipson, Cornell Computational Synthesis Lab: adviser.
\item Jim T\o rreson, University of Oslo: adviser.
\item Juan Zagal, University of Chile: designed and printed robot and provided code for hand tuned gait.
\item Jeff Clune, CCSL Red Couch: collaborated on HyperNEAT implementation/testing.
\item Cooper Bills, Cornell University: assisted with Wii tracker development.
\item Anshumali Srivastava, Cornell University: Teaching Assistant.
\end{itemize}