### 2011-01-20 paper/futureWork.tex

%\section{Future work}

\edit{write this, old section below}

% What are the major shortcomings of your current method? For each 
% shortcoming, propose additions or enhancements that would help overcome it.

There are several directions in which we could continue our
work. First, the error margins for our runs were large, so reducing
them by running more trials would lead to more conclusive data. It
would also be insightful to run our algorithms on different motion
models.  We suspect that our choice of motion model influenced the
results greatly, as even random choices in the space produced gaits
that moved a significant fraction of the time (> 5\%).  It would be
interesting to see how learning methods would perform using a model
that included a much higher percentage of unproductive gaits. We also
intend to experiment further with SVM regression and evolutionary
algorithms/HyperNEAT.  Some parameter vectors resulted in the robot
turning, as opposed to it walking long distances. This could be an
interesting learning goal in future projects. To these ends, we
propose the following additions and enhancements:

\begin{itemize}
\item More runs and/or longer runs
\item Different motion representations
\item Better tuning of SVM regression
\item Evolutionary algorithms/HyperNEAT\cite{clune}
\item Learning how to turn
\end{itemize}