### 2011-01-20 paper/paperOutline.tex

\subsection{Outline of Sections}

\edit{write this}

The remainder of the paper is organized as follows. In \secref{foo}, we
introduce,  \secref{bar} 3 describes.  Finally, we describe future work
in \secref{future}