### 2011-01-20 paper/problemDefinition.tex
Wrote most of Problem Definition

\section{Problem definition}
% Precisely define the problem you are addressing (i.e. formally specify
% the inputs and outputs). Elaborate on why this is an interesting and
% important problem.

\edit{write this.  Quadratot report section below}

We are testing several different learning methods to design a
parametrized gait for a quadruped robot from the Cornell Computational
Synthesis Lab.

The output each of the learning algorithms is a function of time,
$f(t)$, that outputs a vector of commanded motor positions.  This
function is generated using a parametrized motion model, described in
\secref{implement}.

The robot executes these commands and measures its change in location
using the tracking system described in \secref{implement}.  The input
to the learning algorithms is this measured displacement, which the
algorithms attempt to maximize. This displacement is measured for each
gait over a constant length run, usually 12 seconds.

A comparison and evaluation of the many different methods available
for optimizing the gait of legged robots will be useful for future
work on this challenging multidimensional control problem.

\edit{Jason writing this now}

The gait learning problem aims to find a \emph{gait} that maximizes
some performance metric.

Mathematically, we define a gait as a function that specifies
commanded motor positions for a robot over time.  Using this
definition, we require one function per motor.  We can write gaits
without feedback --- also called feed forward gaits --- as

\be
x_i = f_i(t)
\ee

\noindent for commanded position $x_i$ for motor $i$.  The
function depends only on time.  Using the same framework, a gait
including measured motor position, force, or other values could be
written as

\be
x_i = f_i(t, z)
\ee

\noindent for some measurement $z$.  Though perhaps obvious, it is
worth noting that feedforward gaits are deterministic (unless they
include random elements), producing the same command pattern each time
they are run.  The output of a gait that includes feedback, on the
other hand, depends on the actual measurements, and noise in these
measurements may result in different $x_i$ values each time they are
run.

For the system evaluated in this paper, we chose to use only
feedforward gaits.  An interesting extension would be to allow gaits
to depend on the measured servo positions, loads, voltage drops, or
other quantities.

The ultimate goal was to design gaits that were as fast as possible,
so to optimize for speed, our performance metric was displacement over
a period of motion lasting approximately 10 seconds.  For details of
exactly how this displacement was measured, see \secref{foo}.