### 2011-01-20 paper/approach.tex
Split approach into three files

\section{Experimental Approach}

\input{learningMethods}

\input{platformDetails}

\input{fitnessEvaluation}