### 2011-01-21 paper/approach.tex
More work on Experimental Approach and Learning Methods

\section{Experimental Approach}

Given the goal of producing a deterministic, feedforward motion
function $g(t)$, we now turn to several important considerations.  \secref{learningMethods} talks about how this function is generated, \secref{platformDetails} discusses the hardware platform on which it is run, and \secref{fitnessEvaluation} details exactly how its fitness is measured.

\input{learningMethods}

\input{platformDetails}

\input{fitnessEvaluation}