### 2011-01-31 paper/learningMethods.tex
Many edits...

\subsection{Parametrized Motion Models}
\seclabel{motionModel}

There are a plethora of methods available for the creation of the
vector motion generating function $g(t)$ introduced in
\secref{problemDefinition}, and we've chosen two for comparison in
this study.  The first method we evaluated was to use a family of
parametrized gaits, described in this section and the next, and the
second method was to use an evolved non-parametric gait, detailed in
\secref{hyperNeatMethod}.

By a \emph{parametrized gait}, we mean a gait produced by a
parametrized function $g(t; \vec{\theta})$. Fixing the parameter
$\vec{\theta}$ yields a deterministic motion function over time.  We
tried several parametrizations on the robot and, upon obtaining
reasonable early success, settled on one particular parametrization.
We dubbed this the \emph{SineModel5}, as it has five parameters and
employs a sine wave as its root pattern.

The five parameters in $\vec{\theta}$ are as follows:

\edit{make sure this text flows into the table}

%\newcommand{\amp}{\ensuremath{\mathrm{amp}}}
\newcommand{\amp}{\ensuremath{\alpha}}

\begin{table}[h!]
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Parameter    & Description               & Range \\
\hline
\hline
\amp         & Amplitude                 & [0, 400] \\
\hline
$\tau$       & Period                    & [.5, 8] \\
\hline
$m_O$        & Outer motor multiplier    & [-2, 2] \\
\hline
$m_F$        & Front motor multiplier    & [-1, 1] \\
\hline
$m_R$        & Right motor multiplier    & [-1, 1] \\
\hline
\end{tabular}
\caption{The parameters used for the \emph{SineModel5} motion model.}
\tablabel{parameters}
\end{center}
\end{table}

Intuitively, SineModel5 starts with 8 identically sine waves of
amplitude $\amp$ and period $\tau$, multiplies the waves for all inner
motors by $m_o$, multiplies the waves for all front motors by $m_F$,
and multiplies the waves for all left motors by $m_R$.  To obtain the
actual motor position commands, these waves are offset by
appropriate fixed constants so that the base position (when the sine
waves are at 0) is approximately a crouch (the position shown in
\figref{robot_close.jpg}).  Finally, in this motion model, the ninth
``hip'' motor is kept at a neutral position.  Thus, the commanded
position for all motors, as a vector function of time, is:

\[
\vec{g}(t) =
\left[ {\begin{array}{c@{ }l@{ }l}
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \           m_F            & + C_I \\ % 0
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O \cdot m_F            & + C_O \\ % 1
\amp \cdot \sin(2\pi t / \tau) & \                                  & + C_I \\ % 2
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O                      & + C_O \\ % 3
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \                     m_R  & + C_I \\ % 4
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O           \cdot m_R  & + C_O \\ % 5
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \           m_F \cdot m_R  & + C_I \\ % 6
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O \cdot m_F \cdot m_R  & + C_O \\ % 7
0                              & \                                  & + C_C \\ % 8
\end{array} } \right]
\]

\noindent Here the fixed constants that determine the crouched
position are $C_I$ for inner motors, $C_O$ for outer motors, and $C_C$
for the center hip motor.  The nine subscripted functions correspond
to the nine labeled motors in \figref{topdown.png}.

\subsection{Learning Methods for Parametrized Motion Models}
\seclabel{learningMethods}

Using the SineModel5 parametrized motion model described in the
previous section, along with the allowable ranges for each of the five
parameters (shown in \tabref{parameters}), the task becomes how to
choose the combination of five parameters that results in the fastest
motion, per the evaluation methods in \secref{fitnessEvaluation}.

If we choose a value for the five dimensional parameter
$\vec{\theta}$, then a given physical trial gives us one measurement
of the fitness of that parameter vector, or $f(\vec{\theta})$.  Two
things make learning difficult.  First, each evaluation of
$f(\vec{\theta})$ is expensive, taking 15-20 seconds of time on
average.  Second, the fitness returned by such evaluations has proved
to be very noisy, with the standard deviation of the noise often being
roughly equivalent to the size of the measurement.

For both of these reasons, an intelligent \emph{learning algorithm} is
needed.
In this context, a learning algorithm is a method for choosing
%\begin{enumerate}
%\item The value for $\vec{\theta}$ for the initial trial, and
the next value of $\vec{\theta}$ to try, given a list of the
$\vec{\theta}$ values tried so far and the fitness $f(\vec{\theta})$
measured for each.
%\end{enumerate}

We evaluated a total of seven subtly or not so subtly different
learning methods for the parametrized motion models.  All methods
needed a way of picking a $\vec{\theta}$ for the first trial.  In
general this is done by selecting a random starting vector within the
allowable domain.  However, as discussed in
\secref{fitnessEvaluation}, we decided to evaluate the different
methods starting at one of several common starting points to reduce
one possible contribution of noise.

%All employed a simple random sampling method for
%requirement (1); that is, all methods picked their initial
%$\vec{\theta}$ value via uniform random sampling within the allowed
%parameter ranges. \edit{make sure this is true. Isn't really, should say that we started with A, B, C}.  Thus, the
%differences in the algorithms was how they selected new $\vec{\theta}$
%values to try from their past experience.

The seven methods used, and their methods for choosing the next
$\vec{\theta}$ are:

%\section{Method}
%\seclabel{method}

% Describe in reasonable detail the algorithm you are using to address
% this problem. A pseudo-code description of the algorithm you are
% using is frequently useful. If it makes sense for your project,
% trace through a concrete example, showing how your algorithm
% processes this example. The example should be complex enough to
% illustrate all of the important aspects of the problem but simple
% enough to be easily understood. If possible, an intuitively
% meaningful example is better than one with meaningless symbols.

%We use several parametrized motion models that command motors to
%positions based on a sine wave, creating a periodic pattern.  While we
%investigated several models, for the bulk of our experiments, we used
%a model whose five parameters are: amplitude, wavelength, scale inner
%vs outer motors, scale left vs right motors, scale back vs front
%motors. Each strategy below attempts to choose the best possible
%parameters for this motion model.

%We implemented and tested 8 different learning strategies.  All
%strategies except for the HyperNEAT method\cite{clune} were
%constrained to pick parameters from within predetermined ranges.

%\edit{JMC:Bullets=narrower column=wasted space. subsubsection instead?}

\begin{itemize}

\item \emph{Random}: This method randomly generates parameter vectors
  in the allowable range for every trial. This strategy was used only
  as baseline for comparison.

\item \emph{Uniform random hill climbing}: This method begins by
  selecting a single random parameter vector.  Subsequent iterations
  generate a neighbor by randomly choosing one parameter to adjust and
  replacing it with a new value chosen with uniform probability in the
  allowable range for that parameter. The neighbor is evaluated by
  running the robot with the newly chosen parameters. If this neighbor
  results in a longer distance walked than the previous best gait, it
  is saved as the new best gait. The process is then repeated, always
  starting with the best gait.

\item \emph{Gaussian random hill climbing}: This method works
  similarly to Uniform random hill climbing, except neighbors are
  generated by adding random Gaussian noise to the current best gait.
  This results in all parameters being changed at once, but the
  resulting vector is always fairly close to the previous best gait.
  We used independently selected noise in each dimension, scaled such
  that the standard deviation of the noise was 5\% of the range of
  that dimension.

\item \emph{N-dimensional policy gradient descent}: In \cite{kohl},
  Kohl and Stone document a method for local gradient ascent for gait
  learning with noisy fitness evaluations, and we have included this
  method in our evaluation.  This strategy explicitly estimates the
  gradient for the objective function. It does this by first
  evaluating \emph{n} randomly generated parameter vectors near the
  initial vector, each dimension of these vectors being perturbed by
  either $-\epsilon$, $0$, or $\epsilon$. Then, for each dimension, it
  groups vectors into three groups: $-\epsilon$, $0$, and $\epsilon$.
  The gradient along this dimension is then estimated as the average
  score for the $\epsilon$ group minus the average score for the
  $-\epsilon$ group. Finally, the method creates a new vector by
  changing all parameters by a fixed-size step in the direction of the
  gradient.

\item \emph{Nelder-Mead simplex method}: The Nelder-Mead simplex
  method \cite{nm} creates an initial simplex with $d+1$ vertices,
  where $d$ is the dimension of the parameter space. The initial
  parameter vector is stored as the first vertex and the other five
  vertices are created by adding to one dimension at a time one tenth
  of the allowable range for that parameter. It then tests the fitness
  of each vertex and based on these fitnesses, it reflects the worst
  point over the centroid in an attempt to improve it.  In general,
  the worst vertex is reflected; however, to prevent cycles and
  becoming stuck in local minima, several other rules are used.  If
  the reflected point is better than the second worst point and worse
  than the best point, then the reflected point replaces the worst. If
  the reflected point is better than the best point, the simplex is
  expanded in the direction of the reflected point. The better of the
  reflected and the expanded point replaces the worst point. If the
  reflected point is worse than the second worst point, then the
  simplex is contracted away from the reflected point. If the
  contracted point is better than the reflected point, the contracted
  point replaces the worst point. If the contracted point is worse
  than the reflected point, the entire simplex is shrunk \cite{nm}.

\item \emph{Linear regression}: To initialize, this method chooses and
  evaluates five random parameter vectors. It then fits a linear model
  from parameter vector to fitness. In a loop, the method chooses and
  evaluates a new parameter vector generated by taking a fixed-size
  step in the direction of the gradient for each parameter, and fits a
  new linear model to all vectors evaluated so far, choosing the model
  to minimize the sum of squared errors.

\item \emph{SVM regression}: Similarly to linear regression, this
  model starts with several random vectors, but this time they are
  chosen in a small neighborhood about some initial random vector.
  These vectors (generally 8) are evaluated, and a support vector
  regression model is fit to the observed fitnesses.  To choose the
  next vector for evaluation, we randomly generate some number
  (typically 100) of vectors in the neighborhood of the best observed
  gait, and select for evaluation the vector with the best predicted
  performance.  We suspected that if we always chose the best
  predicted point out of 100, we may end up progressing along a narrow
  subspace, prohibiting learning of the true local fitness function.
  Put another way, we would always choose exploitation of knowledge
  vs. exploration of the space.  To address this concern, we added a
  parameter dubbed \code{bumpBy} that added noise to the final
  selected point before it was evaluated.

  Such a method naturally has many tunable parameters, and we
  endeavored to select these parameters by tuning the method in
  simulation.  To estimate the performance of the algorithm, we ran it
  against a simulation with a known optimum.  The simulated function
  was in the same five dimensional parameter space, and simply
  returned a fitness determined as the height of a Gaussian with a
  random mean.  The width of the Gaussian in each dimension was 20\%
  of the range of each dimension, and the maximum value at the peak
  was 100.  \figref{svm_sim_results} shows the learning results on
  this simulated model using the ultimately selected SVM parameters.
  Interestingly, a non-zero value of \code{bumpBy} resulted in better
  learning than noise free (exploration free) learning.

  \edit{JMC: This next paragraph sounds like it belongs in Results?}
  Ultimately, however, the version of SVM tuned for simulation still
  did not show competitive performance on the real robot.  We tried
  tuning some parameters on the real robot, but after some amount of
  tuning, the method still exhibited too little exploration and easily
  became stuck in local minima.

\end{itemize}