### 2011-02-05 paper/approach.tex
Slight approach restructuring, platform details ~done, fixed wiimote fig

% Details of:
%  - hardware
%  - gait generation and learning

\section{Experimental Setup}
\seclabel{experimentalSetup}

\secref{platformDetails} discusses the hardware platform used for the
experiments presented in this paper, and \secref{fitnessEvaluation}
describes the method we used to measure a gait's fitness.

% Given the goal of producing a deterministic, feedforward motion
% function $g(t)$, we now turn to several important considerations.
% \secref{learningMethods} talks about how this function is generated,
% \secref{platformDetails} discusses the hardware platform on which it
% is run, and \secref{fitnessEvaluation} details exactly how its
% fitness is measured.

\input{platformDetails}

\input{fitnessEvaluation}

\section{Gait Generation and Learning}
\seclabel{gaitGenLearn}

\input{learningMethods}

\input{hyperNeatMethod}