### 2011-02-05 paper/paperOutline.tex
Added paperOutline

\subsection{Outline of Sections}

The remainder of the paper is organized as follows. In
\secref{problemDefinition} we define more rigorously the problem of
gait learning. \secref{experimentalSetup} describes our experimental
setup, including the hardware we used and our methods for evaluating
fitness of a gait. \secref{gaitGenLearn} discusses the different gait
generation and learning methods we tested, and \secref{results}
presents and discusses performance results.  Finally,
\secref{conclusion} concludes with possible future extensions to this
work.