### 2011-02-09 paper/acknowledgments.tex
lots of editing of the intro/motivation/previouswork, and some in the robot platform

\section{Acknowledgments}

Left blank for anonymity.

% List any people not on the team who helped you with your project: Your
% TA, other people you consulted with or had useful discussions (say in 
% a word or two what they did), people who proofread your report, and 
% any external code you used (libraries etc).
%\begin{itemize}
%\item Hod Lipson, Cornell Computational Synthesis Lab: adviser.
%\item Jim T\o rreson, University of Oslo: adviser.
%\item Juan Zagal, University of Chile: designed and printed robot and provided code for hand tuned gait.
%\item Jeff Clune, CCSL Red Couch: collaborated on HyperNEAT implementation/testing.
%\item Cooper Bills, Cornell University: assisted with Wii tracker development.
%\item Anshumali Srivastava, Cornell University: Teaching Assistant.
%\end{itemize}