### 2011-02-09 paper/approach.tex
Cleaned up main.tex

% DEPRECATED in favor of separate experimentalSetup.tex and gaitGenLearn.tex files

% % Details of:
% %  - hardware
% %  - gait generation and learning
% 
% 
% 
% \section{Experimental Setup}
% \seclabel{experimentalSetup}
% 
% \secref{platformDetails} discusses the hardware platform used for the
% experiments presented in this paper, and \secref{fitnessEvaluation}
% describes the method we used to measure a gait's fitness.
% 
% % Given the goal of producing a deterministic, feedforward motion
% % function $g(t)$, we now turn to several important considerations.
% % \secref{learningMethods} talks about how this function is generated,
% % \secref{platformDetails} discusses the hardware platform on which it
% % is run, and \secref{fitnessEvaluation} details exactly how its
% % fitness is measured.
% 
% \input{platformDetails}
% 
% \input{fitnessEvaluation}
% 
% 
% 
% \section{Gait Generation and Learning}
% \seclabel{gaitGenLearn}
% 
% \input{learningMethods}
% 
% \input{hyperNeatMethod}