### 2011-02-09 paper/experimentalSetup.tex
Fitness cleanup and rewording, shuffled some points between sections

\section{Experimental Setup}
\seclabel{experimentalSetup}

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world, instead of
using a simulator.  In this section we give details of the hardware
setup we used: \secref{platformDetails} discusses the robotic
platform, and \secref{fitnessEvaluation} describes the tracking system
we used to measure each gait's fitness.

% Given the goal of producing a deterministic, feedforward motion
% function $g(t)$, we now turn to several important considerations.
% \secref{learningMethods} talks about how this function is generated,
% \secref{platformDetails} discusses the hardware platform on which it
% is run, and \secref{fitnessEvaluation} details exactly how its
% fitness is measured.

\input{platformDetails}

\input{fitnessEvaluation}