### 2011-02-09 paper/fitnessEvaluation.tex
More fitness, explanation of bias due to sensor area

\subsection{Fitness Evaluation Details}
\seclabel{fitnessEvaluation}

% Points to make:
% - linear interpolation
% - fitness reported in pixels
% - parametrized models
%   - started all on three thetas
%   - plateued at 1/3 of lifetime
% - HyperNEAT
%   - broke things a lot... so guided evolution (or in HN section?)
%   - eventually included ``shimmy''
% - finite sized window led to unfortunate bias
%
% Given earlier section, now have way of evaluating any gait g(t).
%
% Want to measure average speed, so we measure distance displaced over
% a fixed lenght run. Because we're measureing a single point on
% robot, need to start and end in same position (crouch)
% So we linearly interpolate begining (1s) and end (2s).
%
% Fitness is measured and reported directly as euclidian distance of pixels moved: equation
%
% For the parametrized models, the starting location may be important. To be as fair as possible, we picked three random points and starred all models there.
%   generally ran until 1/3 plateau
%
% For HN, couldn't do a,b,c thing, so just did three runs of length <foo>
% Things started breaking, so included shimmy to test brokenness
% penalized runs that broke servos
%
% Finite sized window kinda hurt stuff.

\acmFigg{wiiMote_crop.jpg}{robot_led_crop.jpg}{1}{A Nintendo Wii
  remote provided the location of the robot by tracking the infrared
  LED mounted on the robot's antenna.  The LED position was measured
  in pixels and transmitted from the Wii remote to the robot via
  bluetooth.}

To track the position of the robot and thus determine gait fitness, we
mounted a Nintendo Wii remote on the ceiling and an infrared LED on
top of the robot (\figref{wiiMote_crop.jpg}).  The Wii remote contains
an IR camera that can track and report the position of up to four IR
sources in its image frame.  The resolution of the camera was 1024 by
768 pixels with view angles of about 40$^\circ$ by 30$^\circ$, which
produced a resolution of 1.7mm per pixel when mounted at a height of
2.63m.  At this height, the viewable window on the floor was
approximately 175 x 120 cm.

A separate tracking server, written in Python, ran on the robot and
interfaced with the Wii remote via bluetooth using the CWiid
library\cite{cwiid}.  Our fitness testing code communicated with this
server via socket connection and requested position updates at the
beginning and end of each run.

As mentioned earlier, the metric for evaluating gaits was their
average speed.  In reality this was measured as the distance the robot
moved over a 12 second run.  The Euclidean distance between the
starting position and the ending position, in pixels, was reported
directly as the fitness (save for a tweak for HyperNEAT; see below).

Since only a single point on the robot --- the IR LED --- was measured
for the purposes of computing fitness, it was important that the
position of the IR LED accurately reflect the position of the robot as
a whole.  To enforce this constraint, the robot was always measured
while in the \emph{ready} position (the position shown in
\figref{robot_close.jpg}).  This was done to prevent assigning extra
fitness to, for example, gaits that ended with the robot leaning into
the direction of travel.  While this would mean that a part of the
robot has traveled farther, this extra distance would not generalize
to a longer run.  Thus, we chose not to reward this behavior.

In order to measure start and end position in the same pose, we
linearly interpolated the motion of the robot between the ready
position and the commanded gait, $g(t)$.  As shown in
\figref{linearInterp}, motion for the first second was interpolated
between the ready position and $g(t)$, for the next nine seconds the
robot moved exactly as specified by $g(t)$, and for the final two
seconds its motion was interpolated smoothly back to the ready
position.

\acmFig{linearInterp}{1}{To ensure fair fitness evaluations with as
  little noise as possible, the position of the robot was measured at
  the beginning and end of each run (red circles) in the same
  \emph{ready} pose (\figref{robot_close.jpg}).  Motion was
  interpolated linearly between this stationary pose and the commanded
  gait g(t) for one second at the beginning of the run and two seconds
  at the end, as shown above.}

% finite size
% shimmy

The only human intervention required during most learning trials was
to occasionally move the robot back into the viewable area of the Wii
remote if it had walked outside.  Initially this was a rare
occurrence, as the gaits did not typically produce motion as large as
the size of the window (roughly 175 x 120 cm).  However, as gaits
improved, particularly when using HyperNEAT, the robot began to walk
out of the measurement area a non-negligible fraction of the time.
Whenever it did so, we would discard the trial and repeat it until the
gait was entirely within the window.

While this process guaranteed that we always obtained a measurement
for a given gait before proceeding, it also biased the measurements
downward.  Because the performance of the robot on a given gait varied
from trial to trial, a successful measurement was more likely to be
obtained when the gait happened to perform poorly.  This phenomenon
was negligible at first, but became significantly more pronounced when
evaluating several gaits that produced motion across nearly the entire
area.  For future studies we will likely use a 2x2 array of Wii
remotes to double the length and width of the measurement arena.

%We made one final tweak to the fitness measurement process.  HN..... 1/2