### 2011-02-10 paper/futureWork.tex
Conclusion/Future Work

%\section{Future work}
\seclabel{futureWork}

Because all trials were done in hardware, it was difficult to gather
the thousands of trials that would be necessary to establish a more
rigorous statistical ranking of the methods' performance, or their
asymptotic behavior as longer and longer training runs are allowed.
Thus, one extension of the current work would be to obtain these
thousands more trials.

However, a more effective extension might be to combine trials in
simulation and in hardware, perhaps doing only every tenth or
hundredth trial in hardware.  The simulation would produce the
necessary volume of trials to allow the learning methods to be
effective, and the hardware trials would serve to continuously ground
and refine the simulator.  One could also guide evolution to the most
fertile territory by penalizing gaits that produced large
discrepancies between simulation and reality.

Another natural extension would be to allow gaits with feedback;
i.e. gaits that depend not only on time, but also the measured
position of the legs, voltage drop across the servos, etc.