### 2011-02-10 paper/introduction.tex
intro edits

\section{Introduction and Background}

\acmFig{robot_close.jpg}{1}{The quadruped robot. The translucent parts
  were created with 3D-printing technology. Videos of the gaits
  evolved for this robot can be viewed online~\cite{quadratotWeb}.}

Gaits for walking robots are often designed with some explicit or
implicit goal in mind.  For some applications, the design criteria may
be obvious --- perhaps the robot needs to move as quickly or as
efficiently as possible --- but other times the objective is more complicated,
requiring simultaneous optimization of several desired traits, each
with its own relative importance.  The different combinations of desired traits and the relative weight placed on each can produce drastically different gaits.    For example, Honda's Asimo
\cite{chestnutt2006footstep} and Boston Dynamic's Big Dog \cite{raibert2008bigdog} both require
gaits that are relatively quick, power efficient, and robust to
changing terrain, but they vary widely in the importance placed
on each attribute.  Big Dog's gait is optimized for much more difficult terrain than Asimo's, resulting in a gait of a completely different form.

Once the design goals are decided upon, gaits may be obtained by one
of several methods.  Gaits may be designed manually by an expert, as
for Asimo and Big Dog \cite{chestnutt2006footstep, raibert2008bigdog}.
Alternately, they may be learned through repeated trial and
evaluation.  Learned gaits offer several advantages over manually
designed gaits.  Automatically learning gaits can save valuable
engineering time and can allow customization of gaits to a particular
robot and its unique actuators.  Most importantly, in many cases
learned gaits can outperform manually designed
gaits~\cite{hornby2005autonomous, valsalam2008modular}.

In this paper we compare the performance of two different methods of
learning gaits: parameterized gaits optimized with six different
learning methods, and gaits generated by evolving neural networks with
the HyperNEAT generative encoding~\cite{stanley2009hypercube}. While
some of these methods, such as HyperNEAT, have been tested in
simulation~\cite{clune2009evolving, clune2011performance}, we
investigate how they perform when evolving on a physical robot
(\figref{robot_close.jpg}).

Previous work has shown that quadruped gaits perform better when they
are \emph{regular} (i.e.\ when the legs are
coordinated)~\cite{clune2009evolving,
  clune2011performance,valsalam2008modular}. For example, HyperNEAT
produced fast, natural gaits in part because its bias towards regular
gaits created coordinated movements that outperformed gaits evolved by
an encoding not biased towards regularity~\cite{clune2009evolving,
  clune2011performance}. One of the motivations of this paper is to
investigate whether any learning method biased towards regularity
would perform well at producing quadruped gaits, or whether
HyperNEAT's high performance is due to additional factors, such as its
abstraction of biological development (described below). We test this
hypothesis by comparing HyperNEAT to six local search algorithms with
a parametrization biased toward regularity.

A final motivation of this problem is to simply evolve effective gaits
for a physical robot. Because HyperNEAT performed well in simulation,
it is interesting to test whether it can produce a fast gait for a
physical robot. It is additionally interesting to test how more
traditional gait optimization techniques compete with evolutionary
algorithms when evolving in hardware.

% Put back in for reference... some good text here I think

% Applications of walking robots often call for the ability to walk as 
% quickly, efficiently, or with as little power as possible.  Often 
% these optimizations are done manually by an expert who designs and 
% tweaks a gait specifically for a given objective.  Other groups have 
% used learning methods to generate gaits optimized for some metric. 
% Approaches differ in their starting assumptions, some essentially 
% tweaking the parameters of a hand-tuned model \cite{chernova}, others 
% exploring a reasonably compact parameter space \cite{kohl}, and still 
% others beginning with few assumptions besides periodicity 
% \cite{zykov}. 
%  
% We aimed to strike a middle ground between these approaches.  Our 
% motion generator did not rely on hand-tweaked gaits, but it did use 
% parameterized gaits which, by their nature, impose some assumptions on 
% the answers produced.  We then used machine learning to design gaits 
% for a quadruped robot with these models.  This paper presents a 
% comparison of the different learning methods implemented.  Most 
% methods created walks that are several times faster than the original 
% hand-tuned gait.  We invite readers with short attention spans to view 
% a video of some of our results online here: 
%  
% 
% %\url{http://www.youtube.com/watch?v=ODoiOj9DdGg} 
% \texttt{http://www.youtube.com/watch?v=ODoiOj9DdGg}