### 2011-04-06 paper_ecal/main.tex
Converted paper to ECAL format, still need to fix section references

%
% Paper for ECAL conference, formatted per alifexi.sty
% JBY 4/6/11
%

\documentclass[letterpaper]{article}
\usepackage{natbib,alifexi}   %% The order is important
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{color}
%\usepackage{ulem}
\usepackage{ownstyles}
%\usepackage{hyperref}

\graphicspath{{../figures/}}

\title{Generating gaits for physical quadruped robots: evolved neural networks vs. local parameterized search}
\author {Jason Yosinski$^{1}$,
Diana Hidalgo$^{1}$,
Sarah Nguyen$^{1}$,
Jeff Clune$^{1}$,
Juan Cristobal Zagal$^{2}$,
\and Hod Lipson$^{1}$\\
\mbox{} \\
$^1$ Cornell University, 239 Upson Hall, Ithaca, NY  14853, USA\\
$^2$ University of Chile, Beauchef 850, Santiago 8370448, Chile\\
jy495@cornell.edu}

\begin{document}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%

% Section 1
\input{introduction}
\input{relatedWork}
\input{paperOutline}

% Section 2
\input{problemDefinition}

% Section 3
\section{Experimental Setup}
\seclabel{experimentalSetup}

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world, instead of
using a simulator.  In this section we give details of the hardware
setup we used: \secref{platformDetails} discusses the robotic
platform, and \secref{fitnessEvaluation} describes the tracking system
we used to measure each gait's fitness.

\input{platformDetails}
\input{fitnessEvaluation}

% Section 4
\section{Gait Generation and Learning}
\seclabel{gaitGenLearn}
\input{learningMethods}
\input{hyperNeatMethod}

% Section 5
\input{results}

% Section 6
\input{conclusion}
\input{futureWork}

% Section 7
\input{acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%

\footnotesize
\bibliography{references.bib}
\bibliographystyle{apalike}

\end{document}