### 2011-04-12 paper_ecal/platformDetails.tex

\subsection{Platform details}
\seclabel{platformDetails}

%\seclabel{implement}

% Describe how you implemented your system and how you structured it. 
% This should give an overview of the system, not a detailed 
% documentation of the code. The documentation of the code is part of 
% the code you hand in. You might want to comment on high-level design 
% decisions that you made. Also explain how you obtained your
% data and any pre-processing you did to it.

\acmFig{topdown2.jpg}{1}{(a) Top-down perspective of the robot with
  the nine joints and associated servos labeled. (b) The robot in a
  flat pose with the hip joint centered. (c,d,e) Various views of a
  pose in which the hip joint is rotated.}

The quadruped robot used in this study was assembled from parts
printed on the Objet Connex 500 3-D Printing System. It weighs 1.88 kg
with the on-board computer and measures approximately 38 centimeters
from leg to opposite leg in the crouch position depicted in
\figref{robot_whitebg.jpg}. The robot is actuated by 9 AX-12+
Dynamixel servos: one inner joint and one outer joint servo in each of
the four legs, and one servo at the center ``hip'' joint.  This final
unique servo allows the two halves of the robot to rotate with respect
to each other. \figref{topdown2.jpg} shows this unique motion, as well
as the positions and numerical designations of all nine servos.  Each
servo could be commanded to a position in the range [0,~1023],
corresponding to a physical range [-120$^{\circ}$,~+120$^{\circ}$].

All of the computation for gait learning, fitness evaluation, and
robot control was performed on the compact, on-board CompuLab Fit-PC2,
running Ubuntu Linux 10.10. 
The most computationally expensive portion of code was HyperNEAT, but even
this took less than one second to run on the on-board computer. Thus,
we chose not to offload any computation.
All gait generation, learning, and fitness
evaluation code, except HyperNEAT, was written in Python and is
available on our website \cite{quadratotWeb}.  HyperNEAT is written in
C++.  We controlled the servos with the \code{pydynamixel} library
\cite{pydynamixel}, sending commanded positions at 40Hz.  The robot
connected to a wireless network on boot, which enabled us to control
it via SSH.

To run a gait on the robot, one needed only to provide a \emph{gait
  function} --- any Python function that accepts a single input, time
starting at 0, and outputs a list of nine commanded positions, one for
each servo.  To safeguard against limb collision with the robot body,
the robot control code cropped the commands to a safe range.  This
range was [150,~770] for the inner leg servos, [30,~680] for the outer
leg servos, and [392,~623] for the center hip servo.

%The robot was set up to run when given any giat

%is run using a given motion model, including, if desired, 
%smooth interpolation over time for the beginning and end of the run.

%The servos are prevented from being commanded to a point outside their
%normal range (0 - 1023) as well as beyond points where limbs would collide