### 2011-04-13 paper_ecal/relatedWork.tex

\subsection{Related Work}

Various machine learning techniques have proved to be effective at
generating gaits for legged robots. Kohl and Stone presented a policy
gradient reinforcement learning approach for generating a fast walk on
legged robots\citep{kohl}, which we implemented for comparison. Others
have evolved gaits for legged robots, producing competitive
results~\citep{chernova2005evolutionary, hornby2005autonomous, zykov,
  clune2009evolving, clune2011performance, clune2009hybrid,
  clune2009sensitivity, tellez2006evolving, valsalam2008modular}. In
fact, an evolved gait was used in the first commercially-available
version of Sony's AIBO robot~\citep{hornby2005autonomous}. Except for
work with HyperNEAT~\citep{clune2009evolving, clune2011performance,
  clune2009hybrid, clune2009sensitivity}, the previous evolutionary
approaches have helped evolution exploit the regularity of the problem
by manually decomposing the task. Experimenters have
to choose which legs should be coordinated, or otherwise facilitate
the coordination of motion. Part of the motivation of this paper
is to see if the space of regularities can be explored automatically,
which has previously performed well~\citep{valsalam2008modular}, and to
perform a direct comparison on the same robot to HyperNEAT.