### 2011-04-13 paper_ecal/experimentalSetup.tex
More edits, 75% done :)

\section{Experimental Setup}
\seclabel{experimentalSetup}

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world instead of
using a simulator.  In this section we give details of the hardware
setup we used.  First we discuss the robotic platform on which gaits
were run, and then we describe the tracking system used to quantify
the fitness of those gaits.

%: \secref{platformDetails} discusses the robotic
%platform, and \secref{fitnessEvaluation} describes the tracking system
%we used to measure each gait's fitness.

% Given the goal of producing a deterministic, feedforward motion
% function $g(t)$, we now turn to several important considerations.
% \secref{learningMethods} talks about how this function is generated,
% \secref{platformDetails} discusses the hardware platform on which it
% is run, and \secref{fitnessEvaluation} details exactly how its
% fitness is measured.

%\input{platformDetails}

%\input{fitnessEvaluation}