### 2011-04-14 paper_ecal/experimentalSetup.tex
more forging

\section{Experimental Setup}
\seclabel{experimentalSetup}

We next describe the details of the experimental system.

%: \secref{platformDetails} discusses the robotic
%platform, and \secref{fitnessEvaluation} describes the tracking system
%we used to measure each gait's fitness.

% Given the goal of producing a deterministic, feedforward motion
% function $g(t)$, we now turn to several important considerations.
% \secref{learningMethods} talks about how this function is generated,
% \secref{platformDetails} discusses the hardware platform on which it
% is run, and \secref{fitnessEvaluation} details exactly how its
% fitness is measured.

%\input{platformDetails}

%\input{fitnessEvaluation}