### 2011-04-14 paper_ecal/futureWork.tex

%\section{Future work}
\seclabel{futureWork}

Because all trials were done in hardware, it was difficult to gather
the many trials that would be necessary to properly rank the methods statistically.
One direction for future work could be to obtain many more trials. However, a more effective extension might be to combine frequent trials in
simulation with infrequent trials in hardware~\citep{bongard}.  The simulation would produce the
necessary volume of trials to allow the learning methods to be
effective, and the hardware trials would serve to continuously ground
and refine the simulator.  One could also guide evolution to the most
fertile territory by penalizing gaits that produced large
discrepancies between simulation and reality~\citep{koos2010crossing}. Another extension would be to allow gaits that sensed the position of the robot and other variables to enable the robot to adjust to its physical state, instead of providing an open-loop sequence of motor commands. All of these approaches would likely improve the quality of automatically generated gaits for legged robots, which will hasten the day that humanity can benefit from their vast potential.