### 2011-04-18 paper_gecco_2/abstract.tex
Cut 8 page version to 2 pages. 1:42:49 :)

Creating gaits for legged robots is an important task to enable robots
to access rugged terrain, yet designing such gaits by hand is a
challenging and time-consuming process. In this paper we investigate
various algorithms for automating the creation of quadruped
gaits. Because many robots do not have accurate simulators, we test
gait learning algorithms entirely on a physical robot.  We compare the
performance of two classes of learning gaits: locally searching
parameterized motion models and evolving artificial neural networks
with the HyperNEAT generative encoding.  All parameter search methods
outperform a manually-designed reference gait, but HyperNEAT performs
better still, producing gaits nearly 9 times faster than the reference
gait.

%Overall, this
%research illuminates various approaches to evolving quadruped gaits
%directly in hardware.

%Learning approaches differ in their starting assumptions, some
%tweaking the parameters of a hand-tuned model, others exploring a
%reasonably compact parameter space, and still others beginning with
%few assumptions besides periodicity.