### 2011-04-18 paper_gecco_2/futureWork.tex

%\section{Future work}
\seclabel{futureWork}

Because all trials were done in hardware, it was difficult to gather
the many trials that would be necessary to properly rank the methods
statistically.  One direction for future work could be to obtain many
more trials. However, a more effective extension might be to combine
frequent trials in simulation with infrequent trials in
hardware.
%The simulation would produce the necessary
%volume of trials to allow the learning methods to be effective, and
%the hardware trials would serve to continuously ground and refine the
%simulator.  One could also guide evolution to the most fertile
%territory by penalizing gaits that produced large discrepancies
%between simulation and reality~\cite{koos2010crossing}.
%Another
%natural extension would be to allow gaits that sensed the position of
%the robot and other variables to enable the robot to adjust to its
%physical state, instead of providing an open-loop sequence of motor
%commands.
%All of these approaches would likely improve the quality of
%automatically generated gaits for legged robots, which will hasten the
%day that humanity can benefit from their vast potential.