### 2011-07-13 poster_gecco/poster.tex
Working on poster, 70% finished.

\documentclass[portrait,final]{baposter}

\usepackage{times}
\usepackage{calc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{ownstyles}
\usepackage{bm}

\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{multicol}

\usepackage{pgfbaselayers}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\usepackage{helvet}
%\usepackage{bookman}
\usepackage{palatino}

\newcommand{\captionfont}{\footnotesize}

\renewcommand\refname{}
%\renewcommand\bibname{New Title2}

\selectcolormodel{cmyk}

\graphicspath{{../figures/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Some math symbols used in the text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Format 
\newcommand{\Matrix}[1]{\begin{bmatrix} #1 \end{bmatrix}}
\newcommand{\Vector}[1]{\Matrix{#1}}
\newcommand*{\SET}[1]  {\ensuremath{\mathcal{#1}}}
\newcommand*{\MAT}[1]  {\ensuremath{\mathbf{#1}}}
\newcommand*{\VEC}[1]  {\ensuremath{\bm{#1}}}
\newcommand*{\CONST}[1]{\ensuremath{\mathit{#1}}}
\newcommand*{\norm}[1]{\mathopen\| #1 \mathclose\|}% use instead of $\|x\|$
\newcommand*{\abs}[1]{\mathopen| #1 \mathclose|}% use instead of $\|x\|$
\newcommand*{\absLR}[1]{\left| #1 \right|}% use instead of $\|x\|$

\def\norm#1{\mathopen\| #1 \mathclose\|}% use instead of $\|x\|$
\newcommand{\normLR}[1]{\left\| #1 \right\|}% use instead of $\|x\|$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multicol Settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setlength{\columnsep}{0.7em}
\setlength{\columnseprule}{0mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Save space in lists. Use this after the opening of the list
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\compresslist}{%
\setlength{\itemsep}{1pt}%
\setlength{\parskip}{0pt}%
\setlength{\parsep}{0pt}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Here starts the poster
%%%---------------------------------------------------------------------------
%%% Format it to your taste with the options
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\typeout{Poster Starts}
\background{
  \begin{tikzpicture}[remember picture,overlay]%
    \draw (current page.north west)+(-2em,-0em) node[anchor=north west] {\hspace{-2em}\includegraphics[height=1.1\textheight]{silhouettes_background}};
  \end{tikzpicture}%
}
\definecolor{silver}{cmyk}{0,0,0,0.3}
\definecolor{blue1}{cmyk}{0.3,0.05,0.05,0.0}
\definecolor{darkblue1}{cmyk}{1.0,0.5,0.5,0.0}
\definecolor{black}{cmyk}{0,0,0.0,1.0}
\definecolor{darkYellow}{cmyk}{0,0,1.0,0.5}
\definecolor{darkSilver}{cmyk}{0,0,0,0.1}

\definecolor{lightorange}{rgb}{1.0, 0.9, 0.8}
\definecolor{lighteryellow}{cmyk}{0,0,0.1,0.0}
\definecolor{lightestyellow}{cmyk}{0,0,0.05,0.0}
\begin{poster}{
  columns = 3,
  % Show grid to help with alignment
  grid=no,
  % Column spacing
  colspacing=1em,
  % Color style
  bgColorOne=white,
  bgColorTwo=white,
  borderColor=darkblue1,
  headerColorOne=blue1,
  headerColorTwo=darkblue1,
  headerFontColor=black,
  boxColorOne=white,
  boxColorTwo=white,
  % Format of textbox
  textborder=roundedleft,
  % Format of text header
  eyecatcher=no,
  headerborder=open,
  headerheight=0.09\textheight,
  headershape=roundedright,
  headershade=plain,
  headerfont=\normalsize\textsf, %Sans Serif
  boxshade=plain,
%  background=shade-tb,
  background=plain,
  linewidth=2pt
  }
  % Eye Catcher
  {} % No eye catcher for this poster. If an eye catcher is present, the title is centered between eye-catcher and logo.
  % Title
  {\sf %Sans Serif
  %\bf% Serif
  \vspace{0em}{\huge Generating Gaits for Physical Quadruped Robots:} \\{\LARGE Evolved Neural Networks Vs. Local Parameterized Search}}
  % Authors
  {\sf %Sans Serif
  % Serif
  \vspace{.3em}\\{\normalsize Jason Yosinski, Jeff Clune, Diana Hidalgo, Sarah Nguyen, Juan Cristobal Zagal, Hod Lipson}\\\hspace{0.3em}
  \code{\footnotesize \{jy495, jeffclune, djh283, smn64\}@cornell.edu, jczagal@ing.uchile.cl, hod.lipson@cornell.edu}\hspace{3em}
  }
  % University logo
  {{\begin{minipage}{.25\textwidth}
    \hfill
    \includegraphics[height=5.0em]{cornellLogo}
  \end{minipage}}
  }

  \tikzstyle{light shaded}=[top color=baposterBGtwo!30!white,bottom color=baposterBGone!30!white,shading=axis,shading angle=30]

  % Width of left inset image
     \newlength{\leftimgwidth}
     \setlength{\leftimgwidth}{0.78em+8.0em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Now define the boxes that make up the poster
%%%---------------------------------------------------------------------------
%%% Each box has a name and can be placed absolutely or relatively.
%%% The only inconvenience is that you can only specify a relative position 
%%% towards an already declared box. So if you have a box attached to the 
%%% bottom, one to the top and a third one which should be in between, you 
%%% have to specify the top and bottom boxes before you specify the middle 
%%% box.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %
    % A coloured circle useful as a bullet with an adjustably strong filling
    \newcommand{\colouredcircle}[1]{%
      \tikz{\useasboundingbox (-0.2em,-0.32em) rectangle(0.2em,0.32em); \draw[draw=black,fill=baposterBGone!80!black!#1!white,line width=0.03em] (0,0) circle(0.18em);}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Abstract}{name=abstract,column=0,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \input{abstract}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Platform}{name=platform,column=0,below=abstract}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    %\pwfig[l]{robot_whitebg.jpg}{.5}{The quadruped robot for which
    %  gaits were evolved. The translucent parts were produced by a 3D
    %  printer. Videos of the gaits can be viewed at
    %  http://bit.ly/geccogait}
    %
    %\pwfig[r]{topdown2.jpg}{.3}{(a) Top-down perspective of the robot
    %  with the nine joints and associated servos labeled. (b) The
    %  robot in a flat pose with the hip joint centered. (c,d,e)
    %  Various views of a pose in which the hip joint is rotated.}

    \pfigg{robot_whitebg.jpg}{topdown2.jpg}{.49}{The 3D printed quadruped robot. Videos of the gaits can be viewed at
      http://bit.ly/geccogait}

The quadruped robot in this study was assembled from off-the-shelf components and parts
printed on the Objet Connex 500 3-D Printing System. It weighs 1.88 kg
with the on-board computer and measures approximately 38 centimeters
from leg to opposite leg in the crouch position depicted above.
The robot is actuated by 9 AX-12+
Dynamixel servos: one inner joint and one outer joint servo in each of
the four legs, and one servo at the center ``hip'' joint.  This final
unique servo allows the two halves of the robot to rotate with respect
to each other.

All of the computation for gait learning, fitness evaluation, and
robot control was performed on the compact, on-board CompuLab Fit-PC2,
running Ubuntu Linux 10.10.
All gait generation, learning, and fitness
evaluation code, except HyperNEAT, was written in Python and is
available on our website (http://bit.ly/geccogait).

  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Experimental Setup}{name=experimentalSetup,column=1}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\pfigg{wiiMote_crop.jpg}{robot_led_crop.jpg}{1}{A Nintendo Wii
%  remote provided the location of the robot by tracking the infrared
%  LED mounted on the robot's antenna.  The position was measured
%  in pixels and transmitted from the Wii remote to the robot via
%  bluetooth.}

\pwfigg{wiiMote_crop.jpg}{robot_led_crop.jpg}{.6}{A Nintendo Wii
  remote tracked an IR LED on the robot, transmitting the (x,y)
  location via bluetooth.}

%To track the position of the robot and thus determine gait fitness, we
%mounted a Nintendo Wii remote on the ceiling and an infrared LED on
%top of the robot.  The Wii remote contains
%an IR camera that tracks and reports the position of IR sources.
The
resolution of the camera was 1024 by 768 pixels with view angles of
about 40$^\circ$ by 30$^\circ$, which produced a resolution of 1.7mm
per pixel when mounted at a height of 2.63m.  At this height, the
viewable window on the floor was approximately 175 x 120 cm.

\pwfig[r]{linearInterp}{.6}{Motion was interpolated linearly between a
  stationary pose and the commanded gait $g(t)$.}

\vspace{.5em}

The fitness was the Euclidian distance the robot
moved during a 12-second run on flat terrain.
The HyperNEAT gaits stressed the motors
more than the other gaits, so to encourage gaits that did not tax the
motors, we penalized gaits that caused the servos to stop
responding by a factor of 1/2.

The robot was always measured while in the \emph{ready} position
(shown in the first figure) to ensure fair fitness measurements. As
shown below the instantaneous robot limb configuration during the
first and last portions of the evaluation was an interpolation between
the initial ready position and $g(t)$; during the rest of the
evaluation, the robot followed the commanded gait exactly.

  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Parameterized Gaits}{name=parameterizedGaits,column=1,span=1,below=experimentalSetup}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Parameterized gaits were produced by the following gait function $g(t; \vec{\theta})$:

%\newcommand{\amp}{\ensuremath{\mathrm{amp}}}
\newcommand{\amp}{\ensuremath{\alpha}}

{
\footnotesize
\[
\vec{g}(t) =
\left[ {\begin{array}{c@{ }c@{ }c@{ }l@{ }l}
\amp \cdot \sin(2\pi t / \tau) & \             & \ \cdot \ m_F & \            & + C_I \\ % 0
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O & \ \cdot \ m_F & \            & + C_O \\ % 1
\amp \cdot \sin(2\pi t / \tau) & \             & \             & \            & + C_I \\ % 2
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O & \             & \            & + C_O \\ % 3
\amp \cdot \sin(2\pi t / \tau) & \             & \             & \ \cdot \ m_R  & + C_I \\ % 4
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O & \             & \ \cdot \ m_R  & + C_O \\ % 5
\amp \cdot \sin(2\pi t / \tau) & \             & \ \cdot \ m_F & \ \cdot \ m_R  & + C_I \\ % 6
\amp \cdot \sin(2\pi t / \tau) & \ \cdot \ m_O & \ \cdot \ m_F & \ \cdot \ m_R  & + C_O \\ % 7
0                              & \             & \             & \            & + C_C \\ % 8
\end{array} } \right]
\]
}

\noindent with parameters in the following ranges:

{
\footnotesize
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Parameters        &                           &       \\
in $\vec{\theta}$ & Description               & Range \\
\hline
\hline
\amp         & Amplitude                 & [0, 400] \\
\hline
$\tau$       & Period                    & [.5, 8] \\
\hline
$m_O$        & Outer-motor multiplier    & [-2, 2] \\
\hline
$m_F$        & Front-motor multiplier    & [-1, 1] \\
\hline
$m_R$        & Right-motor multiplier    & [-1, 1] \\
\hline
\end{tabular}
\tablabel{parameters}
\label{tab:params}
\end{center}
}

Intuitively, SineModel5 starts with 8 identical sine waves of
amplitude $\amp$ and period $\tau$, multiplies the waves for all outer
motors by $m_O$, multiplies the waves for all front motors by $m_F$,
and multiplies the waves for all right motors by $m_R$.  To obtain the
actual motor position commands, these waves are offset by
fixed constants ($C_O = 40$ for outer
motors, $C_I = 800$ for inner motors, and $C_C = 512$ for the center hip motor) so that the base position (when the sine
waves are at 0) is approximately a crouch.
To keep the size of the model search space as
small as possible, we decided to keep the ninth (center) motor at a fixed neutral
position.

\vspace{.5em}

The \textbf{learning methods} used to search the parameterized gait
space were: \textbf{Random}, \textbf{Uniform random hill climbing},
\textbf{Gaussian random hill climbing}, \textbf{N-dimensional policy
  gradient ascent}, \textbf{Nelder-Mead simplex method}, \textbf{Linear
  regression}, and \textbf{SVM regression}.

  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{HyperNEAT Gaits}{name=hyperneatGaits,column=1,span=1}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Gaits generated via HyperNEAT were compared to the parameterized gaits.

\pfig{hyperneatExplanation.png}{1}{HyperNEAT produces ANNs from
  CPPNs \cite{clune2011performance}.}

HyperNEAT is an indirect encoding for evolving artificial neural
networks (ANNs) that is inspired by the way natural organisms
develop~\cite{stanley2009hypercube}. It evolves Compositional Pattern
Producing Networks (CPPNs)~\cite{stanley2007compositional}, each of
which is a genome that encodes an ANN
\pwfig{SpiderANN.jpg}{.6}{The ANN configuration used to generate gaits
  \cite{clune2011performance}.}\label{hyperneatLayout}
phenotype~\cite{stanley2009hypercube}. Each CPPN is itself a directed
graph, where the nodes in the graph are mathematical functions, such as sine or
Gaussian. The nature of these functions can facilitate the evolution
of properties such as symmetry (e.g.\ a Gaussian function) and repetition (e.g.\ a sine
function)~\cite{stanley2009hypercube, stanley2007compositional}.

For quadruped locomotion, HyperNEAT has previously produced gaits with
front-back, left-right, and four-way
symmetries~\cite{clune2009evolving, clune2011performance}.

\pfig{neat_110115_211410_00000_002_filt_zoom}{1}{Joint angles for a
  high-performance HyperNEAT gait.}

  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Results}{name=results,column=2,span=1,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    By training on data collected in the lab, we were able to predict
    quadrotor drift quite accurately.  The below images show the
    actual versus predicted drift, including the directly predicted
    derivatives as well as the inferred (integrated) position.

    %\begin{multicols}{2}

      %\begin{center}
      %  \includegraphics[width=0.9\linewidth]{dec17_1_DataLog-1292574599_07_int}
      %\end{center}

    %\end{multicols}\vspace{-1em}

    %%%%%%%%%%%%%%%%%%%%%%
    \mbox{\hspace{0.3\linewidth}\rule{0.4\linewidth}{1pt}\hspace{0.3\linewidth}}\\

    Drift predictions under normal conditions, trained on data
    collected in similar conditions.
    \vspace{-.7em}

    \begin{multicols}{3}

      \begin{tabular}{ccc}
        \includegraphics[width=.9\linewidth]{robot_whitebg.jpg}
        &
        \includegraphics[width=.9\linewidth]{robot_whitebg.jpg}
        &
        \includegraphics[width=.9\linewidth]{robot_whitebg.jpg}
        \\

        \capt{Predicted versus actual x drift at each timestep in
          pixels per second.} & \capt{Predicted versus actual y drift
          at each timestep in pixels per second.} & \capt{Actual
          position versus position inferred from integrated
          predictions}
      \end{tabular}\\

    \end{multicols}\vspace{-1em}

    %%%%%%%%%%%%%%%%%%%%%%
    \mbox{\hspace{0.3\linewidth}\rule{0.4\linewidth}{1pt}\hspace{0.3\linewidth}}\\

    Predictions under artificially windy conditions, trained on data
    collected under normal conditions as well as windy conditions.
    \vspace{-.7em}

    \begin{multicols}{3}

      \begin{tabular}{ccc}
        \includegraphics[width=.9\linewidth]{robot_whitebg.jpg}
        &
        \includegraphics[width=.9\linewidth]{robot_whitebg.jpg}
        &
        \includegraphics[width=.9\linewidth]{robot_whitebg.jpg}
        \\

        \capt{Predicted vs. actual x drift} & \capt{Predicted
          vs. actual y drift} & \capt{Integrated position}
      \end{tabular}\\

      \end{multicols}\vspace{-1em}

    %%%%%%%%%%%%%%%%%%%%%%
    \mbox{\hspace{0.3\linewidth}\rule{0.4\linewidth}{1pt}\hspace{0.3\linewidth}}\\

    Although our algorithm was able to accurately predict drift, we
    are still working on getting the prediction code to be fast enough
    to control the quadrotor in live flight.  The SVM regression code
    runs quickly, but we're currently facing difficulties with delays
    due to I/O as well as a suboptimal configuration.\\

  }%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Conclusion}{name=conclusion,column=2,span=1,below=results}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    We are making strong progress towards predicting and removing
    quadrotor drift.  The drift is predicted fairly accurately via a
    60 dimensional sensor data vector from the quadrotor, and we hope
    to soon demonstrate a controller enhanced by the predicted drift. \vspace{.5em}
  }%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Abbreviated References}{name=references,column=0,above=bottom}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \vspace{-1em}

    \footnotesize
    \bibliography{references.bib}
    \bibliographystyle{is-plain}
  }%
\end{poster}%
%
\end{document}