### 2012-05-30 paper_aimsa/relatedWork.tex

\section{Related Work}

Other attempts to solve it