### 2012-05-31 paper_aimsa/experimentalSetup.tex
5/31/2012

\section{Experimental Setup}

hardware details, ...
Setup:
The quadruped robot used was assembled from parts purchased online and parts printed by the Objet Connex 500 3-D Printing System. The robot actuation system consists of 5 AX-18+ Dynamixel servos and 4 AX-12+ Dynamixel servos: one inner joint with one AX-18+ servo and one outer joint with AX-12+servo in each of the four legs, and one AX-18+ servo at the center. To avoid the formerly reported AX-18+ servos are used in this robot because of their stronger actuation power than that of AX-12+. Each servo could be set to a position in the range [0, 1023] by using PyDynamixel library, corresponding to a physical range [-120, +120]. Also, to prevent collisions with the robot body, the control module filter out the commands to a safe range. This range was [-85, +60 ] for the inner leg servos, [-113 , +39 ] for the outer leg servos, and [-28 , +28] for the center hip servo. In the studies of this paper, tethered cables powered both the computer and the servos. It measures approximately 39.5 centimeters from leg to opposite leg in the crouch position showed in [pic].

Our performance metric was the displacement over the evaluation period of 12 seconds for each. Same as Jason et al described in [xx], the displacement was measured using a Wii remote that was placed on the ceiling. Different from the original model described in [xx], the quadruped robot was equipped with a three-infrared-LED cluster on top rather than just one. The reason for this setup is that when fierce gaits were executed, the Wii remote loses tracking of the robot position due to the limited visible angle of a single LED.  These three LEDs were place tightly together to act as one signal emitter. Each LED was tilted outwards in order to maximize the visible range. [pic]

A separate tracking server ran on the robot PC interacted with the Wii remote via bluetooth by using the CWiid library.  A corresponding client communicates with this server via a socket connection. Our fitness evaluation code talks with the Wii remote by using this client and updates the position from beginning to end during each run.

The metric for evaluating gaits was the Euclidian distance the robot traveled during a 12-second run on flat surface. Previous work done in [xx] and [xx1(k)] suggested that extremely fierce are not viable in general. These gaits tend to either overburden the servos or flip the robot. Thus, RL_POWER(?) were tested after carefully cropping out the extremely fierce gaits each time. (As described earlier, each leg has two joints, inner joint j1 and outer j2. The position of one leg is determined by the sum of the values of j1 and j2.

Extremely fierce gaits usually have relatively small values of g. Thus cropping works by mapping the wild gaits onto the boundary of a triangular space over the two dimensional space of j1 and j2. [pic of j1 j2 space] [g = j1 + j2] caption