### 2012-06-01 paper_aimsa/conclusion.tex
finally

\section{Conclusion}

Conclusion and Future Work
We have presented a new reinforcement-learning-based algorithm for optimizing a quadrupedal gait for linear speed. We implemented and tested six learning strategies for parameterized gaits and compared them to gaits produced by neural networks evolved with the HyperNEAT generative encoding. Though both methods resulted in an improvement over the robot’s previous naïve gaits, based on the statistics collected, RL_PoWER has a more elegant and consistent performance.

Based on previous studies on evolving gait on quadruped robots[xx], RL_PoWER produced higher performing gaits than all of the parameterized methods. Its best-performing gait traveled XX body lengths per minute.  HyperNEAT, showed in the previous studies, achieved the competitive best performance with RL_PoWER. HyperNEAT gaits

Though over 900 trials have been made to investigate the applicability of RL_PoWER to quadruped robots. It is difficult to gather the enough trials that would be necessary to properly rank the methods. One direction for future work could be to obtain many more trials. But due to the physical limitations, obtaining  One solution to this is simulation. Because of the low cost of simulation, it would produce the necessary volume of trials to allow the learning methods to be effective, and the hardware trials would serve to continuously ground and refine the simulator. [KYRREG XX]

One guess led by this study is that for feedback oriented tasks, reinforcement learning methods are more fit in natural. Despite the complexities of HyperNEAT, a simpler algorithm on the code level delivered better performance in general.  Also, evolvable spline interpolation is shown to be simple and representationally powerful at the same time. Can serve as a general representation for various other learning problems.