### 2012-06-01 paper_aimsa/introduction.tex
?

\section{Introduction}

\figp{quadratotWhiteBooties}{.75}{}

Various learning algorithms have been tested to be effective for
legged robots. These studies tested the viability of applying RL to legged robot gaits learning. Competitive performance has
been verified on methods such as HyperNEAT \cite{yosinski2011evolving-robot-gaits} and
GA \cite{chernova2004an-evolutionary-approach-to-gait} and other methods \cite{hornby2005autonomous-evolution-of-dynamic} \cite{zykov2004evolving-dynamic-gaits} \cite{tellez2006evolving-the-walking-behaviour} \cite{valsalam2008modular-neuroevolution-for-multilegged}. Despite the excellent performance of these algorithms, a major
part of achieving good performance relies in tuning the parameters for
the evolutionary algorithms \cite{kormushev2011bipedal-walking-energya}. Here we present a
different way for learning gaits using reinforcement learning and expectation maximization. In our
experiment, it turned out this method achieved competitive performance
with

\subsection{Problem Definition}
The gait-learning problem is defined to find a
gait that maximizes some specific metric. Mathematically, we define a
gait as a function that specifies a vector of commanded motor
positions for a robot over time. Gaits without feedback --- also
called open-loop gaits --- can be defined as

\be\vec{x} = g(t)\ee

According to its definition, open-loop gaits are deterministic. One
particular gait should behave exactly the same when it is run from
trial to trial. However, the actual robot motion and fitness measured
will vary due to the errors and uncertainty of the real world physics. In our
trials, the gaits generated were sent to the robot and
executed in an open loop manner. We will present the performance comparison between
RL POWER and HyperNEAT, the latter of which have the best in details. The metric used will be
described later.