### 2012-06-02 paper_aimsa/relatedWork.tex

\section{Related Work}
Many attempts have been made on robot gaits learning using machine learning algorithms, producing competitive results (Chernova and Veloso, 2005; Hornby et al., 2005; Zykov et al., 2004; Clune et al., 2009a, 2011, 2009b,c; T´ellez et al., 2006; Valsalam and Miikkulainen, 2008). In fact, Jason et al. tested and compared six different algorithms for quadruped robot gait learning in their studies. HyperNEAT, a generative encoding neural network algorithm, achieved the best performance among all six methods. In Kormushev et al. (2011)'s bipedal robot energy reduction research, a reinforcment learning method was used to optimize the performance. The results from Kormushev's research showed this algorithm's poetential for walking problems. This reinforcement learning method still needs more tests for further explorations.