### 2012-06-08 paper_aimsa/main.tex

\documentclass[letterpaper]{article}
\usepackage{natbib,alife13mod}
\usepackage{ownstyles}

\graphicspath{
  {../figures/}
  {../figures/aracna_legs/}
}

\title{RL-PoWER, Evolving Policy Parameterization with Reinforcement Learning and its Application to Quadrupedal Walking Optimization}
\author{Haocheng Shen, Jason Yosinski, Petar Kormushev, \and Hod Lipson  \\
  \mbox{}\\
Cornell University, 239 Upson Hall, Ithaca, NY 14853 \\
\texttt{hs454@cornell.edu, yosinski@cs.cornell.edu}}

\begin{document}
\maketitle

%%%%%%%%%%%%%
% Abstract
%%%%%%%%%%%%%

\begin{abstract}
\input{abstract}
\end{abstract}

%%%%%%%%%%%%%
% Main Paper
%%%%%%%%%%%%%

\input{introduction}

\input{relatedWork}

\input{methods}

\input{experimentalSetup}

\input{results}

\input{conclusion}

\input{acknowledgements}

%%%%%%%%%%%%%
% References
%%%%%%%%%%%%%

\footnotesize
\bibliographystyle{apalike}
\bibliography{references}
\end{document}