### 2012-06-18 paper_aimsa/relatedWork.tex
Next round of edits.

\section{Related Work}

Many attempts have been made on robot gaits learning using machine
learning algorithms, producing competitive results \edit{Change these
  all to citations} (Chernova and Veloso, 2005; Hornby et al., 2005;
Zykov et al., 2004; Clune et al., 2009a, 2011, 2009b,c; T´ellez et
al., 2006; Valsalam and Miikkulainen, 2008). In fact, Jason et
al. tested and compared six different algorithms for quadruped robot
gait learning in their studies. HyperNEAT, a generative encoding
neural network algorithm, achieved the best performance among all six
methods. In previous bipedal robot energy reduction
research \citep{kormushev2011bipedal-walking-energy}, a reinforcement learning method was used to optimize the
performance. The results from this research showed the
algorithm's potential for walking problems. This reinforcement
learning method still needs more tests for further explorations.