### 2012-06-25 paper_aimsa/report/abstract.tex

% A brief summary of the report. This is not an introduction -
% it should be complete. Include 1-2 sentences for each of the items
% below, up to and including conclusions.
An evolvable spline representation has been tested on bipedal walking energy reduction. 
Combined with expectation-maximization, we proposed a new reinforcement way for quadupedal robot gait learning for walking speed maximization. 
We compare results with HyperNEAT, a previously tested efficient 
method for this problem. The proposed new method is tested to be better on average and best gait than the HyperNEAT. In the meanwhile, we tried to solve some existing issue with the testing platform---QuadraTot. New methods was added to the existing 
robot control library and fixed over-fierce action issue. 
%Because the fastest learned walk
%was not significantly faster than the fastest randomly generated walk,
%we conjecture that the motion representation for the robot is more
%integral to forward speed than the learning algorithm.

Keywords: Platform Improvement, Learning Control, Walking Robots, Gait Learning, Quadruped