### 2012-08-14 paper_sean/bibliography.tex

\bibitem {bongard:lipson}
Bongard, J., Zykov, V., and Lipson, H.:
Resilient Machines Through Continuous Self-Modeling.
Science. 314(5802):1118-1121 (2006)
%
\bibitem {valsalam:mii}
Valsalam, V. and Miikkulainen, R.:
Modular Neuroevolution for Multilegged Locomotion.
In: Proceedings of the Genetic and Evolutionary Computation Conference, pp. 265--272 (2008)
%
\bibitem {kohl:stone}
Kohl, N. and Stone, P.:
Machine Learning for Fast Quadrupedal Locomotion.
In: The Ninetheenth National Conference on Artificial Intelligence (AAAI 2004), pp. 611--616. San Francisco (2004)
%
\bibitem {clune1}
Clune, J., Beckmann, B., Ofria, C., and Pennock, R.:
Evolving Coordinated Quadruped Gaits with the HyperNEAT Generative Encoding.
In: Proceedings of the IEEE Congress on Evolutionary Computation, pp. 2764--2771 (2009)
%
\bibitem {clune3}
Clune, J., Ofria, C., and Pennock, R.:
The Sensitivity of HyperNEAT to Different Geometric Representations of a Problem.
In: Proceedings of the Genetic and Evolutionary Commuptation Conference, pp. 675--682 (2009)
%
\bibitem {stanley1}
Stanley, K. O., D'Ambrosio, D. B., and Gauci, J.:
A Hypercube-Based Encoding for Evolving Large-Scale Neural Networks.
Artificial Life. 15(2), 185-212 (2009)
%
\bibitem {clune2}
Clune, J., Stanley, K. O., Pennock, R., and Ofria, C.:
On the Performance of Indirect Encoding Across the Continuum of Regularity.
IEEE Transactions on Evolutionary Computation. 15, 346-367 (2011)
%
\bibitem {yos:clune}
Yosinski, J., Clune, J., Hidalgo, D., Nguyen, S., Zagal, J. C.:
Evolving Robot Gaits in Hardware: the Hyperneat Generative Encoding vs. Parameter Optimization.
In: Proceedings of the 20th European Conference on Artificial Life, pp. 11--18 (2011)
%
\bibitem {glette}
Glette, K., Kalus, G., Zagal, J. C., and Torresen, J.:
Evolution of Locomotion in a Simulated Quadruped Robot and Transferral to Reality.
In: Proceedings of the Seventeenth International Symposium on Aritifical Life and Robotics, (2012)
%
\bibitem {haocheng}
Shen, H., Yosinski, J., Kormushev, P., Caldwell, D. G., Lipson, H.:
Learning Fast Quadruped Robot Gaits with the RL PoWER Spline Parameterization.
In: AIMSA 2012 Workshop on Advances in Robot Learning and Human-Robot Interaction, (2012)
%
\bibitem {robotis}
Robotis:
User's Manual, Dynamixel AX-12.
URL: http://www.trossenrobotics.com/images/productdownloads/AX-12(English).pdf
%
\bibitem {clune:lipson}
Clune, J., Lipson, H.:
Evolving Three-Dimensional Objects with a Generative Encoding Inspired by Developmental Biology.
In: Proceedings of the European Conference on Artificial Life, pp. 144--148 (2011)
%
\bibitem {stanley2}
Stanley, K. O.:
Compositional Pattern Producing Networks: A Novel Abstraction of Development.
Genetic Programming and Evolvable Matter. 8(2), 131--152 (2007)
%
\bibitem {stanley3}
Stanley, K. O.:
Exploiting Regularity Without Development.
In: Proceedings of the AAAI Fall Symposium on Developmental Systems, (2006)
%
\bibitem {stanley4}
Stanley, K. O., Miikkulainen, R.:
Evolving Neural Networks Through Augmenting Topologies.
Evolutionary Computation, 10(2):99-127 (2002)
%
\bibitem {jakobi}
Jakobi, N., Husbands, P., and Harvey, I.:
Noise and the Reality Gap: The Use of Simulation in Evolutionary Robotics.
In: Advances in Artificial Life: Proceedigns of the 3rd European Conference on Artificial Life, pp. 704--720, Springer-Verlang, (1995)
%%
\bibitem {koos1}
Koos, S., Mouret, J.-B., Doncieux, S.:
Crossing the Reality Gap in Evolutionary Robotics by Promoting Transferable Controllers.
In: Proceedings of the 12th Annual Conference on Genetic and Evoutionary Computation, pp. 119--126 (2010)
%
\bibitem {carroll}
Carroll, S.:
Endless Forms Most Beautiful: The New Science of Evo Devo and the Making of the Animal Kingdom.
Norton, New York (2005)
%
\bibitem {bongard}
Bongard, J.:
Synthesizing Physically-Realistic Environmental Models from Robot Explorations.
In: Proceedings of the 9th European Confrence on Advances in Artificial Life, pp. 806--815 (2007)
%
\bibitem {zagal}
Zagal, J., Ruiz-del Solar, J., and Vallejos, P.:
Back to Reality: Crossing the Reality Gap in Evolutionary Robotics.
In: Proceedings of IAV 2004, the 5th IFAC Symposium on Intelligent Autonomous Vehicles, (2004)
%
\bibitem {koos2}
Koos, S., Mouret, J.-B., Doncieux, S.:
The Transferability Approach: Crossing the Reality Gap in Evolutionary Robotics.
IEEE Transactions on Evolutionary Computation, (2011)