### 2012-10-30 paper_sean/conclusion.tex
jeff done editing...pending feedback

% !TEX root =main.tex

With a simulator, HyperNEAT evolved the fasted gait yet recorded for the QuadraTot robot, outperforming nine machine learning algorithms from three previous publications~\cite{yos:clune,glette,haocheng}, including an improvement of 54.3\% in simulation and 5.4\% in reality over the previous best QuadraTot gait by Glette et al. These results provide an important demonstration that the HyperNEAT generative encoding can evolve state-of-the-art results for challenging engineering problems, in this case evolving gaits for a legged robot. The results further reaffirm the benefits of using simulators when solving real-world challenges with evolutionary algorithms. Our results additionally confirm--with a different robot and simulator--previous work that has shown that HyperNEAT is an effective encoding for automatically evolving coordinated, regular gaits in simulation~\cite{clune2009evolving,clune2011performance,yos:clune}. That HyperNEAT outperformed the encoding hand-designed by Glette et al. shows that HyperNEAT can even outperform evolutionary algorithms manually designed to incorporate human domain knowledge regarding which regularities are thought to be beneficial for a problem, demonstrating that automatically discovering regularities can be a superior approach to specifying them even for problems that are relatively well understood.