### 2012-10-30 paper_sean/conclusion.tex
minor

% !TEX root =main.tex

With a simulator, HyperNEAT evolved the fasted gait yet recorded for the QuadraTot robot, outperforming nine machine learning algorithms from three previous publications~\cite{yos:clune,glette,haocheng}, including an improvement of 54.3\% in simulation and 5.4\% in reality over the previous best QuadraTot gait by Glette et al. These results provide an important demonstration that the HyperNEAT generative encoding can evolve state-of-the-art results for challenging engineering problems, in this case evolving gaits for a legged robot. The results further reaffirm the benefits of simulators when solving real-world challenges with evolutionary algorithms. Our results additionally confirm---with a different robot and simulator---previous work that has shown HyperNEAT to be an effective encoding for automatically evolving coordinated, regular gaits in simulation~\cite{clune2009evolving,clune2011performance,yos:clune}. That HyperNEAT outperformed the encoding hand-designed by Glette et al. shows that HyperNEAT can outperform even evolutionary algorithms manually designed to incorporate human domain knowledge regarding which regularities are thought to be beneficial, demonstrating that automatically discovering regularities can be a superior approach to specifying them, even on problems that are relatively well understood.